---
title: "Demonstration of foofactors new capabilities"
author: "Elijah Willie"
output:
  github_document:
    toc: yes
---

# Introduction
Hello my name is `Elijah Willie` and welcome to a modified version of the `foofactor` package. In this document, I will attempt to show the new functionalities of this package.



# Installtion

Below is how one would go about installing this package

```{r}
suppressMessages(library(dplyr))
library(devtools)
# devtools::install_github("STAT545-UBC-students/hw07-ecool50")
```

# Functions in foofactor
Factors are a very useful type of variable in R, but they can also drive you nuts. Especially the "stealth factor" that you think of as character.

Can we soften some of their sharp edges?

## fbind
Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work.  Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

## freq_out

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```


## isFactor

One of the new functionalities is the `isFactor` function. This function checks to see if a factor could be represented as a string. If the number of levels of the factor is equal to the length of the factor, then it is most likely a character rather than a factor.


```{r}
#show the case where it is indeed a factor
x <- c('X', 'X', 'Y')
x_new <- factor(x)
y <- c('W', 'X')
y_new <- factor(y)
isFactor(x_new)
isFactor(y_new)
```

We see that for the first case it returns `False` and for the second case it returns `True`

## Factor_Reorder

The `Factor_Reorder` function takes in a factor and returns the levels of the factor reordered. The reordering is done in descending order.
```{r}
#create a factor variable
x <- c('X', 'X', 'Y', 'Q', 'P')
x_new <- factor(x)
#get the levesl of x
x_levels <- levels(x_new)

#now reorder the factor levels
x_reord <- Factor_Reorder(x_new)

#print the results
print("Original factor levels")
x_levels

print("Reordered factor levels")
x_reord
  
```

## Factor_Order

The function `Factor_Order` takes in a factor and returns the levels of the factor in the same order as they appear in the dataset.
```{r}
#create a factor variable
x <- c('X', 'X', 'Y', 'Q', 'P')
x_new <- factor(x)

print("the original factor levels")
levels(x_new)

print("After setting the factors as is")
Factor_Order(x_new)
```

We see that the factors are returned as is in the original dataset.

## Read_textFile
Now I will demonstrate reading in a textfile and automatically detecting its delimeter

```{r}
#set the data source
datae_source <- "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt"

#read in the data
df <- Read_textFile(datae_source)

#show it
knitr::kable(head(df, 10))

#get some summary
str(df)

```

Looking at the structure of the data, we can see that both country and continent was read in as factors as desired.

## Write_textFile

Finally, I will demonstrate the writing to text file functionality. 
```{r}
library(gapminder)
#create a new dataframe to write
df_new <- data.frame(gapminder$country, gapminder$lifeExp)

#write it to a textfile
Write_textFile(df_new, name = "Gapminder_cont_lifeExp.txt", delim = "\t")

#read it back to double check
df_gap <- Read_textFile("Gapminder_cont_lifeExp.txt")

#double check it was written correctly
knitr::kable(head(df_gap, 10))
```

Fabulous! everything works!

# Credits

* Some of the work here may have been borrowed and modified from other sources.
* This was purely done to illustrate some ideas
* I take no credit for parts of the work that are not mine
* Let credit be given where it is due.
* The foofactor package was originally written by `Jennifer Bryan` and can be found [here](https://github.com/jennybc/foofactors)
